A Survey of Range-Only SLAM Techniques and Sensors

### 2018-11-12 Paper.tex
Added more recent references
Added more references to the paper from 2016+.

Citation abrv can be seen on the google doc.

Made minor grammer and spelling edits all over the paper. Some sections still seem to need significant rewording/organization.... but will probably be acceptable even without that.

		While the loss of orientation from traditional, vision based landmarks  proves to be a detriment, range only beacons solve two major problems encountered in traditional landmark-based approaches: landmark association and landmark scaling. The landmark association problem occurs when a detected landmark has similar features to previously recorded landmark and is mistaken for the other, generating huge errors in the estimation of both the robot and the landmark. This problem is avoided using beacons, as they have some kind of network address which is broadcast during transmissions, so there is no way to confuse which range is associated with which landmark. The scaling problem occurs when a SLAM algorithm runs for a long time and more and more landmarks are appended to the map. As more landmarks are added, the amount of memory needed to store the positions and distributions of these landmarks increases, as well as computation time. When using beacons, a finite number of landmarks exists, and the total size of the map is known before deployment. In addition, fewer landmarks are needed in beacon-based applications, as most beacons do not require line-of-sight to be measured, unlike visual landmarks. Since the robot can observe the landmarks more often, it doesn't need to add nearly as many to ensure accurate position estimation. %Need a citatation to back that claim up

@@

			Wireless Fidelity (wi-fi)  is the a popular beacon choice as existing networks are already prevalent in many indoor locations and they are generally low cost \cite{Youssef2005} \cite{Sen2013} \cite{Kotaru2015} \cite{Rea2017}. \cite{Youssef2005} presents a system known as Horus with the aim of high accuracy and low computation cost for wireless local area network location determination. \cite{Sen2013} proposes a method named CUPID which extracts information from the physical layer to process information allowing it to avoid the impact of multipathing. \cite{Kotaru2015} presents a design and implementation for its SpotFi system which can be deployed on top of existing wi-fi infrastructure. \cite{Rea2017} designed a new method for filtering range measurment specifically designed with the large noise found in indoor WiFi radios in mind.

@@

	\section{Methods and Their Implementations}
	\label{methods and their implementations}
	The following implementations are broken down into categories based on the general algorithm that is used by the implementation to do RO-SLAM. The concept behind the algorithm is summarized in each category followed by a discussion of implementations that incorporates said algorithm. Each category concludes by examining the problems that current implementations are focusing on and theorizing about possible future directions for research.  
	\subsection{EKF}

@@

	 When used to solve SLAM the map applied to the EKF is a feature based map, meaning it is composed of observable features (landmarks) which can be distinguished between during re-observation \cite{Thrun2002}. This distinction becomes important when examining RO-SLAM where only the range to a landmark, or the range between landmarks, is known. 
	 %This restriction means that landmarks often need to be manufactured 
	 %add rererence here...
	 %, such as the beacons discussed in previous sections, and physically placed in an environment. 
	 In addition, the exact location of the landmarks, or even there initial existence, is not always known at the beginning of runtime. Initialization is a critical element in many algorithms that make use of the EKF. If the initialization is inaccurate then the convergence to a correct estimation can become much more difficult. The EKF has seen a wide variety of implementation ranging from ground based robotics\cite{Djugash2008} \cite{Shue2017}, to unmanned air vehicles \cite{Fabresse2016}, and to autonomous underwater vehicles \cite{Olson2006}. This section will proceed to give a summary of the various ways EKF has been implemented in regards to RO-SLAM.

	%Underwater implementations
	A system for robust range only localization was prosed by \cite{Olson2006}. The system utilized an Autonomous Underwater Vehicle (AUV) and beacons without carefully surveyed locations. The paper notes that one of the majors issue was noise present in the received data. A method for outliers rejection caused by the noise was accomplished via spectral graph partitioning. A voting scheme was implemented similar to a Hugh transform to get approximate beacon locations \cite{Hough1959}. The approximate locations were then incorporated into an EKF. It is noted by the paper that this method is particularly CPU intensive, however they state that since the rate at which data is acquired from a landmark is slow, every 5 or so seconds, that the AUV can afford the expensive computation since there will be an adequate amount of time between readings for the data to process. The methodology was simulated using a dataset and can be seen in the paper.

	%AUV implementations
	\cite{Kurth2003} presents a comparison of how three methods process a collected data set. The methods are a EKF, a sliding batch, and a particle filter. The comparison considers the cross-track error, and along track error. The results state that the particle filter works well when used to acquire an initial pose estimate, and EKF works well with real-time localization. It is suggested that in the future that the two could be switched between when appropriate to improve the overall system.

@@

	\cite{Vallicrosa2015} presents a solution utilizing a Sum of Gaussian (SOG) filter for a single range only beacon. It uses multivariate Gaussian mixtures to represent the probability distribution functions of the robot and beacon positions as Gaussian particles which each contain an EKF for vehicle position, velocity, and estimated beacon position. The paper presented experimentation where the method was used to map a beacon while correcting the navigation of the vehicle.

@@

	An extension to EKF called relative-over parametrized (ROP) EKF is presented by \cite{Djugash2008}. This extension uses specific parameterization to  better represent the range only data utilized in RO-SLAM. ROP EKF operates in polar coordinates as opposed to the Cartesian space that typical EKF operates. The method provided improved results over EKF when sparse range measurement or data association errors were present. \cite{Herranz2014} presents a comparison of the ROP-EKF method presented by \cite{Djugash2008}, to a smoothing and mapping method presented by \cite{Dellaert2006}. Experimentation conducted by \cite{Herranz2014} on and indoor and outdoor environment show an improvement in the results of the smoothing and mapping (SAM) method over ROP EKF.

@@

	\cite{Fabresse2013} presents a solution for the 3D RO-SLAM problem. The solution reduces computational load through reduced spherical parametrization of map feature positions. The paper also proposes a new EKF update method which incorporates a reduced spherical representation of the state vector. In addition the paper describes how to efficiently integrate multi-modal belief of elevations angles and azimuth from a range sensor into the EKF utilizing two independent Gaussian Mixtures \cite{Fabresse2013}. In addition \cite{Fabresse2014} expands upon this by presenting a pre-filtering algorithm to be applied to range measurement before being used in the EKF in order to reduce the effect of noise generated by environmental issue such as multipath. It is then proposed by \cite{Fabresse2015} that the above method can be utilized for decentralized multi-SLAM with range only sensors. This new method would allow for landmarks estimated by a robot to be incorporated into the estimation of another nearby robot, thus improving localization estimation. Based on experimental results obtained by \cite{Fabresse2015} the integration of estimates from multiple robots of the same landmarks provides a benefit to landmark localization over a single robot. In \cite{Fabresse2016} the author also validates and expands upon this method through indoor and outdoor experimentation with an Unmanned Aircraft System (UAS).

@@

In \cite{Torres-Gonzalez2014} the computational complexity is reduced by utilizing the computational capabilities of the nodes or radio  beacons. Here the robot computes the prediction stage of the iteration, sends the information to the beacons where the update process is completed. Hence the computation burden on the robot is reduced. In addition, inter-beacon measurements are factored in to reduce uncertainty even further. In \cite{Torres-Gonzalez2017} the SEIF method is evaluated and compared for 3d SLAM using an octorotor In conclusion this method had better estimates and lesser computational burden than the traditional RO SLAM method using EKF. \cite{Torres-Gonzalez2017a} extends this method by integrating a distributed information-driven measurement allocation tool with the SEIF to allow for dynamic selection of measurements. The end goal being to create a scheme that utilizes robot-beacon and inter-beacon measurements to increase accuracy while maintaining resource-consumption efficiency and scalability. Thus, creating a scalable distributed RO-SLAM scheme with a focus towards resource limited operations.

\subsection{Particle Filters}      %change the references
Particles filters are a category of monte carlo algorithms. They are used to estimate states in partially observable Markov chains as shown in \cite{Doucet2001}. In  early applications, particle filters were used in the field of localization where it was used to estimate a robots pose from sensor data. It is considered better and more robust to conventional estimation methods because of its ability to solve the global localization \cite{Borenstein1996} and the kidnapped robot problem \cite{Engelson1992} which EKF, in comparison, struggles with. At is core, EKF also has higher dimensional computation problems.
% It has also been at the core higher dimensional computational problems.  
%Description
As mentioned in \cite{Thrun2002a}, particle filters outperform EKF in applications involving global localization (the kidnapped robot problem) where the robot is randomly picked up and placed in a unknown environment. 
EKF cannot process negative information, but since particle filters use raw sensor data without any processing they can process negative information\cite{Thrun2002a}.

It works in a similar way to the particle filter initialization procedure mentioned in Section \ref{beacon initialization}, but it uses a weighted average of both the possible landmark and robot positions to converge to the actual robot and landmark locations.

As the number of particles increases the computational complexity of the filter too increases. In order to address this there were several new methods introduced that  exploit the independence among state variables. For example the Rao-Blackwellized particle filters that exploit the independence of the landmarks with the robot path thereby reducing the computation complexity from \cite{Yuan2012}. It also solves the data association problem previously unsolved by EKF. Although this method of localization can be computationally expensive it is preferred whenever the noise involved is non Gaussian.

   \cite{Blanco2008a} uses the particle filter to overcome the delayed initialization issue brought forward by RO SLAM. It uses the Rao-Blackwellized Particle filter to maintain a conditional distribution for each beacon which eventually is transformed into a EKF when the uncertainty becomes sufficiently small. This approach is validated for both simulated and real data sets. This approach is preferred in indoor applications where there is a high probability of multi path signal propagation.

>>>>

		While the loss of orientation from traditional, vision based landmarks  proves to be a detriment, range only beacons solve two major problems encountered in traditional landmark-based approaches: landmark association and landmark scaling. The landmark association problem occurs when a detected landmark has similar features to previously recorded landmark and is mistakenly identified as the incorrect landmark, generating huge errors in the estimation of both the robot and the landmark. This problem can be avoided by many beacons, as often incorporate features such as an constant network address which is broadcast during transmissions, ensuring no way to misidentify which range measurment is associated with which landmark. There are some works that examine scenarios where the range only beacons are indistinguishable, such as in \cite{Jaulin2016} , however the majority of works presented in this paper will consider the beacons to be distinguishable. The scaling problem occurs when a SLAM algorithm runs for a long time and more and more landmarks are appended to the map. As more landmarks are added, the amount of memory needed to store the positions and distributions of these landmarks increases, as well as computation time. When using beacons, a finite number of landmarks exists, and the total size of the map is known before deployment. In addition, fewer landmarks are needed in beacon-based applications, as most beacons do not require line-of-sight to be measured, unlike visual landmarks. Since the robot can observe the landmarks more often, it doesn't need to add nearly as many to ensure accurate position estimation. %Need a citatation to back that claim up

@@

			Wireless Fidelity (WiFi)  is the a popular beacon choice as existing networks are already prevalent in many indoor locations and they are generally low cost \cite{Youssef2005} \cite{Sen2013} \cite{Kotaru2015} \cite{Rea2017}. \cite{Youssef2005} presents a system ,known as Horus, with the aim of high accuracy and low computation cost for wireless local area network location determination. \cite{Sen2013} proposes a method named CUPID which extracts information from the physical layer which is used during processing to alleviate the impact of multipath interference. \cite{Kotaru2015} presents a design and implementation for its SpotFi system which can be deployed on top of existing wi-fi infrastructure. \cite{Rea2017} designed a new method for filtering range measurment specifically designed with the large noise found in indoor WiFi radios in mind.

@@

	\section{Methods and Their Implementations}
	\label{methods and their implementations}
	The following methods implementations are broken down into categories based on the general algorithm that was used to accomplish RO-SLAM. The concept behind the algorithm is summarized in each category followed by a discussion of implementations that incorporates said algorithm. 
	\subsection{EKF}

@@

	When used to solve SLAM the map applied to the EKF is a feature based map, meaning it is composed of observable features (landmarks) which can be distinguished between during re-observation \cite{Thrun2002}. This distinction becomes important when examining RO-SLAM where only the range to a landmark, or the range between landmarks, is known. 
	 %This restriction means that landmarks often need to be manufactured 
	 %add rererence here...
	 %, such as the beacons discussed in previous sections, and physically placed in an environment. 
	In addition, the exact location of the landmarks, or even there initial existence, is not always known at the beginning of runtime. Initialization is a critical element in many algorithms that make use of the EKF. If the initialization is inaccurate then the convergence to a correct estimation can become much more difficult. The EKF has seen a wide variety of implementation ranging from ground based robotics\cite{Djugash2008, Shue2017}, to unmanned air vehicles \cite{Fabresse2016}, and to autonomous underwater vehicles \cite{Olson2006}. This section will proceed to give a summary of the various ways EKF has been implemented in regards to RO-SLAM.

	%Underwater implementations
	A system for robust range only localization was proposed by \cite{Olson2006}. The system utilized an Autonomous Underwater Vehicle (AUV) and beacons without carefully surveyed locations. The paper notes that one of the majors issue was noise present in the received data. A method for outliers rejection caused by the noise was accomplished via spectral graph partitioning. A voting scheme was implemented similar to a Hugh transform to get approximate beacon locations \cite{Hough1959}. The approximate locations were then incorporated into an EKF. It is noted by the paper that this method is particularly CPU intensive, however they state that since the rate at which data is acquired from a landmark is slow, every 5 or so seconds, that the AUV can afford the expensive computation since there will be an adequate amount of time between readings for the data to process. The methodology was simulated using a dataset and can be seen in the paper.

	%AUV implementations
	A comparison of how an EKF, sliding batch, and particle filter evaluate a dataset based on cross and along track error is presented by \cite{Kurth2003}. The results state that the particle filter works well when used to acquire an initial pose estimate, and EKF works well with real-time localization. It is suggested that in the future that the two could be switched between when appropriate to improve the overall system performance.

@@

	An Sum of Gaussian (SOG) filter, which incorporates EKFs, was presented by \cite{Vallicrosa2015} for solving RO-SLAM using range measurements towards only a single beacon. The proposed filter used multivariate Gaussian mixtures to represent the probability distribution functions of the robot and beacon positions as Gaussian particles which each contain an full EKF for vehicle position, velocity, and estimated beacon position. The paper presented experimentation where the method was used to map a beacon while correcting the navigation of the vehicle.

@@

	An extension to EKF called relative-over parametrized (ROP) EKF is presented by \cite{Djugash2008}. This extension uses specific parameterization to  better represent the range only data utilized in RO-SLAM. ROP EKF operates in polar coordinates as opposed to the Cartesian space that typical EKF operates. The method provided improved results over EKF when sparse range measurement or data association errors were present. \cite{Herranz2014} presents a comparison of the ROP-EKF method presented by \cite{Djugash2008}, to a smoothing and mapping method presented by \cite{Dellaert2006}. Experimentation conducted by \cite{Herranz2014} on and indoor and outdoor environment show an improvement in the results of the smoothing and mapping (SAM) method over ROP EKF. The author of \cite{Herranz2014} continues by conducting a comparison of RO-SLAM filtering algorithms utilized indoors via WiFi based sensors. Specifically, EKF, FastSLAM, and SAM are compared in this work \cite{Herranz2016}.

@@

	The joint issue of RO-SLAM and navigation is examined by \cite{Miah2018}. The author explains that they seek to solve the robot localization, motion control, and mapping problem simultaneously in their work. Range measurements are obtained via RSSI from an Xbee radio network and used along with bearing approximations for the measurement model of the EKF. Trajectory information gathered from the mobile robot is used for the prediction phase of the EKF. The map generated by the EKF is then used to plan the navigation of the robot, which influences the control value used to move mobile robot, which is in turn applied to the prediction step of the EKF. Thus, creating a loop which incorporates navigation into the RO-SLAM process.

	\cite{Fabresse2013} presents a solution for the 3D RO-SLAM problem. The solution reduces computational load through reduced spherical parametrization of map feature positions. The paper also proposes a new EKF update method which incorporates a reduced spherical representation of the state vector. In addition the paper describes how to efficiently integrate multi-modal belief of elevations angles and azimuth from a range sensor into the EKF utilizing two independent Gaussian Mixtures \cite{Fabresse2013}. In addition \cite{Fabresse2014} expands upon this by presenting a pre-filtering algorithm to be applied to range measurement before being used in the EKF in order to reduce the effect of noise generated by environmental issue such as multipath noise propagation. It is then proposed by \cite{Fabresse2015} that the above method can be utilized for decentralized multi-SLAM with range only sensors. This new method would allow for landmarks estimated by a robot to be incorporated into the estimation of another nearby robot, thus improving the localization estimation. Based on experimental results obtained by \cite{Fabresse2015} the integration of estimates from multiple robots of the same landmarks provides a benefit to landmark localization over a single robot. In \cite{Fabresse2016} the author also validates and expands upon this method through indoor and outdoor experimentation with an Unmanned Aircraft System (UAS). The author continues in \cite{R.Fabresse2018} by utilizing an undelayed multi-hypothesis EKF approach based on Gaussian Mixture models and a reduced parametrization of the state vector to improve efficiency for 3D spherical based RO-SLAM. The work also incorporates an robot-to-landmark and landmark-to-landmark range only observation model.

@@

	In \cite{Torres-Gonzalez2014} the computational complexity is reduced by utilizing the computational capabilities of the nodes or radio  beacons. Here the robot computes the prediction stage of the iteration, sends the information to the beacons where the update process is completed. Hence the computation burden on the robot is reduced. In addition, inter-beacon measurements are factored in to reduce uncertainty even further. In \cite{Torres-Gonzalez2017} the SEIF method is evaluated and compared for 3d SLAM using an octorotor In conclusion this method had better estimates and lesser computational burden than the traditional RO SLAM method using EKF. \cite{Torres-Gonzalez2017a} extends this method by integrating a distributed information-driven measurement allocation tool with the SEIF to allow for dynamic selection of measurements. The end goal being to create a scheme that utilizes robot-beacon and inter-beacon measurements to increase accuracy while maintaining resource-consumption efficiency and scalability. Thus, creating a scalable distributed RO-SLAM scheme with a focus towards resource limited operations. Further details on and evaluations of the work produced by the author of \cite{Torres-Gonzalez2017,Torres-Gonzalez2017a} can be seen in \cite{Gonzalez2017}.

\subsection{Particle Filters}      %change the references
	Particles filters are a category of monte carlo algorithms. They are used to estimate states in partially observable Markov chains as shown in \cite{Doucet2001}. In  early applications, particle filters were used in the field of localization where it was used to estimate a robots pose from sensor data. It is considered better and more robust to conventional estimation methods because of its ability to solve the global localization \cite{Borenstein1996} and the kidnapped robot problem \cite{Engelson1992} which EKF, in comparison, struggles with. At is core, EKF also has higher dimensional computation problems. 
	% It has also been at the core higher dimensional computational problems.  
	% Description
	As mentioned in \cite{Thrun2002a}, particle filters outperform EKF in applications involving global localization (the kidnapped robot problem) where the robot is randomly picked up and placed in a unknown environment. EKF cannot process negative information, but since particle filters use raw sensor data without any processing they can process negative information\cite{Thrun2002a}. It works in a similar way to the particle filter initialization procedure mentioned in Section \ref{beacon initialization}, but it uses a weighted average of both the possible landmark and robot positions to converge to the actual robot and landmark locations.
	%What in the world is negative information??? ^^^ I don't even k

	As the number of particles increases the computational complexity of the filter increases. In order to address this there were several new methods introduced that  exploit the independence among state variables. For example the Rao-Blackwellized particle filters that exploit the independence of the landmarks with the robot path thereby reducing the computation complexity from \cite{Yuan2012}. It also solves the data association problem previously unsolved by EKF. Although this method of localization can be computationally expensive it is preferred whenever the noise involved is non Gaussian. The work presented by \cite{Samyak2018} utilizes particle filters in an attempt to reduce the impact of indoor multipath propagation error. A series of beacons operating in the ultrawide band frequency are utilized as landmarks with each represented as a full particle filter in the authors implementation. The simulation results provided in \cite{Samyak2018} show an increase in performance of the particle filter method in comparison to several other methods under conditions where multipath propagation is high.

  	\cite{Blanco2008a} uses the particle filter to overcome the delayed initialization issue brought forward in RO-SLAM. It uses the Rao-Blackwellized Particle filter to maintain a conditional distribution for each beacon which eventually is transformed into a EKF when the uncertainty becomes sufficiently small. This approach is validated for both simulated and real data sets. This approach is preferred in indoor applications where there is a high probability of multi path signal propagation.
